---
title: "Spatio-Temporal Analysis Using coefclust"
author: "Junho Lee, Maria Kamenetsky"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to coefclust}
  %\VignetteEngine{knitr::knitr}
  \usepackage[utf8]{inputenc}
---

##Set-Up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

First, load the `coefclust` package and the `geosphere` package.

```{r}
set.seed(1)
# library(coefclust)
# library(geosphere)
```

First, load the `coefclust` package.

```{r}
#TODOD
```

Load in the toy dataset for the spatio-temporal analysis that comes from the package, *SE_FakeData_SpTM*. Explore the dataset using the `str()` function:

```{r}
###################################################################################################
load("../data/SE_FakeData_ST.RData")
str(SE_FakeData_SpTm)
```

This dataset has 11 variables: *State*, *County*, *FIPS* code (as an integer value), *long* (longitude of the centroid for each county), *lat* (latitude of the centroid for each county), 3 covariates (*x1*, *x2*, *x3*), and three responses (*y1*, *y2*, *y2*).


Next, we extract the longitude and latitude into separate objects, `long` and `lat`:

```{r}
long <- SE_FakeData_SpTm$long; lat <- SE_FakeData_SpTm$lat

```





We must first prepare the data into lists. First we prepare the list `yList`, where each element of the list will contain a single series of the three responses:

```{r}
yList <-list()
yList[[1]] <- SE_FakeData_SpTm$y1
yList[[2]] <- SE_FakeData_SpTm$y2
yList[[3]] <- SE_FakeData_SpTm$y3
str(yList)
```

Next, we prepare the list `XList`. Each element of `XList` will have two columns. The first vector is 1 repeated the length of the number of observations. This first column corresponds to the intercept, $\beta_0$. The second column corresponds to the `x` covariate.

```{r}
XList <-list()
XList[[1]] <- cbind(rep(1,length(long)), SE_FakeData_SpTm$x1)
XList[[2]] <- cbind(rep(1,length(long)), SE_FakeData_SpTm$x2)
XList[[3]] <- cbind(rep(1,length(long)), SE_FakeData_SpTm$x3)
str(XList)
```

We start by setting the maximum radius (`MR`) to 300 and the number of simulation (`M`) to 2. In a full simulation, we recommend setting `M=1000`, but have only set it to 2 here for illustration:


```{r}
MR <- 300
#M <- 1000
M <- 2
```

##Detectiong Multiple Clusters Sequentially Using Simultaneous Detection

We use the function `Find.Clusters.SI.ST()` for detectiong multiple clusters sequentially using simultaneous detection. This function will take the list of responses we created (`yList`) as the first argument and the list of covariates (`XList`) as the second argument. The lists of resposnes and covariates are followed by the longitude coordinate, latitude coordinate, the maximum radius for potential clusters `MR`, the number of simulations `M`, we set `overlap=FALSE` to not allow for overlapping clusters, and set `alpha=0.05` as the Type I error rate: 


```{r}
###################################################################################################
### Find Multiple Clusters Sequentially via the Simulatneous Detection ############################
#source("../R/Cluster_Detection_SpTm_Simultaneous.R")
Clusters_simST <- Find.Clusters.SI.ST(yList, XList, long, lat, MR, M, overlap=FALSE, alpha=0.05)
str(Clusters_simST)
Clusters_simST$Clusters
Clusters_simST$Coef
```

The output is an object `Clusters_simST` which is a list of 3 elements:

- `Clusters_simST$Clusters` (first element of list): provides the resulting identified cluster, with information about the center ID, radius, p-value for the cluster, elapsed computation time, and number of observations inside the detected cluster.
- `Clusters_simST$Coef` (second element of list): provides estimates for `beta_0` (intercept) and `beta_1` (slope) for each time period.
- `Clusters_simST$Indicator` (third element of list): is Boolean vector that is `TRUE` if the given center is inside the cluster and `FALSE` if it is not.

To identify the cluster by it’s county name, we can column bind the information from the `SE_FakeData_SpTm` dataset with the identified cluster information:

```{r}
cbind(SE_FakeData_SpTm[Clusters_simST$Clusters[,1],1:2],Clusters_simST$Clusters)
```

##Detecting Multiple Clusters Sequentially Using Two-Stage Detection

```{r}
###################################################################################################
### Find Multiple Clusters Sequentially via the Two--Stage Detection ##############################
source("Cluster_Detection_SpTm_TwoStage.R")
Clusters_ts1N <- Find.Clusters.TS.ST1(yList, XList, long, lat, MR, M, overlap=FALSE, alpha=0.05)
Clusters_ts2N <- Find.Clusters.TS.ST2(yList, XList, long, lat, MR, M, Cls1st=Clusters_ts1N, overlap=FALSE, alpha=0.05)
### With the Bonferroni correction
Clusters_ts1B <- Find.Clusters.TS.ST1(yList, XList, long, lat, MR, M, overlap=FALSE, alpha=(0.05/2))
Clusters_ts2B <- Find.Clusters.TS.ST2(yList, XList, long, lat, MR, M, Cls1st=Clusters_ts1B, overlap=FALSE, alpha=(0.05/2))

###################################################################################################
Clusters_sim$Clusters
cbind(SE_FakeData_SpTm[Clusters_sim$Clusters[,1],1:2],Clusters_sim$Clusters)
Clusters_sim$Coef

Clusters_ts2N$Clusters
cbind(SE_FakeData_SpTm[Clusters_ts2N$Clusters[,1],1:2],Clusters_ts2N$Clusters)
Clusters_ts2N$Coef

Clusters_ts2B$Clusters
cbind(SE_FakeData_SpTm[Clusters_ts2B$Clusters[,1],1:2],Clusters_ts2B$Clusters)
Clusters_ts2B$Coef


```

