---
title: "Introduction to coefclust"
author: "Junho Lee, Maria Kamenetsky"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to coefclust}
  %\VignetteEngine{knitr::knitr}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

First, load the `coefclust` package.

```{r}
library(coefclust)
```

Next load the toy dataset that comes with the package, *SoutheastFakeData*:

```{r}
data("SE_FakeData")
str(SE_FakeData)
```

From the `str()` function, we see that the toy dataset has 7 variables: *State*, *County*, *FIPS* code (as an integer value), *long* - the longitude of the centroid for each county, *lat* - the latitude of the centroid for each county, *x*, and *y* - the vector of responses.

Next, set initial `x` and `y` inputs, the longitude and latitude, maximum radius (`MR`) in km, and number of simulations (`M`). `y` will be equivalent to the response (`y`) in the dataset `SE_FakeData`. For `X`, we must first column bind a vector of 1's for the intercept.

In a full simulation, we recommend setting `M=1000`. However for illustration, we set `M=2` here.

```{r}
y <- SE_FakeData$y
X <- cbind(rep(1,length(y)), SE_FakeData$x)
head(X)
long <- SE_FakeData$long; lat <- SE_FakeData$lat
MR <- 300
M <- 2
```



## Find Multiple Clusters Sequentially via the Simultaneous Detection 

To find multiple clusters sequentially via simultaneous detection, we use the function `Find.Clusters.Simul()`. This function takes as arguments the response vector, `y`, the design matrix `X` we created above, the longitude and latitude coordinates for the centroid for each county, `MR` which is the maximum radius for a potential clusters, `M` the number of simulations, we set `overlap=TRUE` to allow for overlapping clusters, and set `alpha=0.05` as the Type I error rate.

```{r}
Clusters_sim <- Find.Clusters.Simul(y, X, long, lat, MR, M, overlap=TRUE, alpha=0.05)
str(Clusters_sim)
Clusters_sim$Clusters
Clusters_sim$Coef
```

The output is object `Clusters_sim`, which is a list of three elements. 

- `Clusters_sim$Clusters` (first element of list): provides the resulting identified cluster, with information about the center ID, radius, p-value, elapsed computation time, and number of observations in the cluster.
- `Clusters_sim$Coef` (second element of list): provides estimates for `beta_0` (intercept) and `beta_1`.
- `Clusters_sim$clsL`: Boolean vector of TRUE/FALSE indicating locations identified to be inside the cluster (TRUE) or outside of the cluster (FALSE).

To identify the cluster by it's county name, we can column bind the information from the `SE_FakeData` dataset with the identified cluster information:

```{r}
cbind(SE_FakeData[Clusters_sim$Clusters[,1],1:2],Clusters_sim$Clusters)

```


## Find Multiple Clusters Sequentially via the Two-Stage Detection 

To find multiple clusters sequentially via two-stage detection, we use the functions `Find.Clusters.TStg1()` (stage 1) and `Find.Clusters.TStg2()` (stage 2). 

`Find.Clusters.TStg1()` takes as arguments the response vector, `y`, the design matrix `X` we created above, the longitude and latitude coordinates for the centroid for each county, `MR` which is the maximum radius for a potential clusters, `M` the number of simulations, we set `overlap=TRUE` to allow for overlapping clusters, and set `alpha=0.05` as the Type I error rate. The object created by this function (`Clusters_ts1`) becomes an input for the stage 2 function, `Find.Clusters.TStg2()`.

`Find.Clusters.TStg2()` takes as arguments `y`, the design matrix `X` we created above, the longitude and latitude coordinates for the centroid for each county, `MR` which is the maximum radius for a potential clusters, `M` the number of simulations, `Cls1st` which is the output of the first stage and we specify as `Clusters_ts1`, we set `overlap=TRUE` to allow for overlapping clusters, and set `alpha=0.05` as the Type I error rate.


```{r}
#Perform two-stage detection
Clusters_ts1 <- Find.Clusters.TStg1(y, X, long, lat, MR, M, overlap=TRUE, alpha=0.05)
Clusters_ts2 <- Find.Clusters.TStg2(y, X, long, lat, MR, M, Clusters_ts1, overlap=TRUE, alpha=0.05)

#Explore identified clusters (via two-stage detection)
Clusters_ts2$Clusters
Clusters_ts2$Coef

```

The final output of the two-stage detection is in the object `Clusters_ts2`, which is a list of 3 elements:

- `Clusters_sim$Clusters` (first element of list): provides the resulting identified cluster, with information about the center ID, radius, p-value, elapsed computation time, and number of observations in the cluster.
- `Clusters_sim$Coef` (second element of list): provides estimates for `beta_0` (intercept) and `beta_1`.
- `Clusters_sim$clsL`: Boolean vector of TRUE/FALSE indicating locations identified to be inside the cluster (TRUE) or outside of the cluster (FALSE).



To identify the cluster by it's county name, we can column bind the information from the `SE_FakeData` dataset with the identified cluster information:


```{r}
cbind(SE_FakeData[Clusters_ts2$Clusters[,1],1:2],Clusters_ts2$Clusters)

```


